Thesis meeting

### 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
saving gbif wrangling, trying to compile pdf for thesis meeting and edits to preliminary plots

#load data
dat_MA_0<- read_csv("metafor7march.csv")
regions <- read_csv("gbif_regiondata.csv")
```

@@

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) %>%
  filter(resp_units != "C")

@@

### Figure 3.
```{r, message=FALSE}
# boxplots of how fluctuation range influences SMD
ggplot(normalized, aes(x=as.factor(flux_range), y=yi))+
  geom_boxplot()+
  theme_bw()+
  ggtitle("SMD across flux_range")

@@

### Figure 6.
```{r, message=FALSE}
#looking at yi across studies with the same temperature range (10 C)
ggplot(common_range, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
#random effects model within subset data
fig12 <- rma.mv(yi, vi, data=common_range,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig12
#simple linear model looking at how flux_range affects yi within subset data
simple12<-lm(yi~flux_range, data =common_range)
summary(simple12)
```

@@

### Figure 7.
```{r, message=FALSE}
# boxplots of how mean temperature influences SMD
ggplot(normalized, aes(x=as.factor(mean_temp_constant), y=yi))+
  geom_boxplot()+
  theme_bw()+
  ggtitle("SMD across mean temperature")

@@

### Figure 10.
```{r, message=FALSE}
#boxplots of SMD across different genuses colored by body size
ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
  geom_boxplot()+
  theme_minimal()+
  ggtitle("SMD across genuses colored by body size")+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45,hjust = 1), 
        legend.position = "bottom")
#random effects model including genus as a random variable
fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~genus,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig11
#simple linear model looking at how genus affects yi
simple11<-lm(yi~genus, data =dat_MA_ES)
summary(simple11)
```

@@

```{r, message=FALSE}
#plot across studies with common response variable body mass
ggplot(common_unit_g, aes(y=yi, color=as.factor(flux_range), x = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
#random effects model within subset data
fig14 <- rma.mv(yi, vi, data=common_unit_g,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig14
#simple linear model looking at how flux_range affects yi within subset data
simple14<-lm(yi~study_id, data =common_unit_g)
summary(simple14)
```

>>>>

#load data
dat_MA_0<- read_csv("metafor.csv")
regions <- read_csv("gbif_regiondata.csv")
```

@@

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) %>%
  filter(resp_units != "C") %>%
  filter(study_id != "piccau2017")

@@

### Figure 3.
```{r, message=FALSE}
# boxplots of how fluctuation range influences SMD
ggplot(normalized, aes(x=flux_range, y=yi))+
  geom_point()+
  theme_bw()+
  geom_smooth(method="lm", formula = y~x)+
  ggtitle("SMD across flux_range")

@@

### Figure 6.
```{r, message=FALSE}
#looking at yi across studies with the same temperature range (10 C)
ggplot(common_range, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
#random effects model within subset data
fig12 <- rma.mv(yi, vi, data=common_range,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig12
#simple linear model looking at how flux_range affects yi within subset data
simple12<-lm(yi~flux_range, data =common_range)
summary(simple12)
```

@@

### Figure 7.
```{r, message=FALSE}
# scatterplot of how mean temperature influences SMD
ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
  geom_point()+
  theme_bw()+
  ggtitle("SMD across mean temperature")

@@

### Figure 10.
```{r, message=FALSE}
#boxplots of SMD across different genuses colored by body size
ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
  geom_point()+
  theme_minimal()+
  ggtitle("SMD across genuses colored by body size")+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45,hjust = 1), 
        legend.position = "bottom")
#random effects model including genus as a random variable
fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~size,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig11
#simple linear model looking at how genus affects yi
simple11<-lm(yi~genus, data =dat_MA_ES)
summary(simple11)
```

@@

___________________________________
Leftover/excluded plots

```{r, message=FALSE, results = 'hide'}
#plot across studies with common response variable body mass
ggplot(common_unit_g, aes(y=yi, color=as.factor(flux_range), x = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
#random effects model within subset data
fig14 <- rma.mv(yi, vi, data=common_unit_g,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig14
#simple linear model looking at how flux_range affects yi within subset data
simple14<-lm(yi~study_id, data =common_unit_g)
summary(simple14)
```

### 2021-03-08 thesis_meeting7march.Rmd
saving gbif wrangling, trying to compile pdf for thesis meeting and edits to preliminary plots

- S. enterica was the only organism with no data on GBIF and was therefor leftout

```{r}
#extracting occurences for each species as available in gbif
c_reevesii <- occ_search(scientificName = "Chinemys reevesii")
a <- c_reevesii$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

x_arvicola <- occ_search(scientificName = "Xylotrechus arvicola")
b <- x_arvicola$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

g_firmus <- occ_search(scientificName = "Gryllus firmus")
c <- g_firmus$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

l_peronii <- occ_search(scientificName = "Limnodynastes	peronii")
d <- l_peronii$data %>% 
  as.data.frame() %>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

l_tasmaniensis <- occ_search(scientificName = "Limnodynastes	tasmaniensis")
e <- l_tasmaniensis$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

d_simulans <- occ_search(scientificName = "Drosophila simulans")
f <- d_simulans$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

r_prolixus <- occ_search(scientificName = "Rhodnius	prolixus")
g <- r_prolixus$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

c_latirostris <- occ_search(scientificName = "Caiman latirostris")
h <- c_latirostris$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

d_melanogaster <- occ_search(scientificName = "Drosophila melanogaster")
i <- d_melanogaster$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

p_chinensis <- occ_search(scientificName = "Plestiodon chinensis")
j <- p_chinensis$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

#no country column 
e_coli <- occ_search(scientificName = "Escherichia 	coli")
k <- e_coli$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

#zero observations 
s_enetrica <- occ_search(scientificName = "Salmonella	enetrica")$data %>%
  as.data.frame()

d_birchii <- occ_search(scientificName = "Drosophila birchii")
l <- d_birchii$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()
#binding all the data together
gbif_data <- bind_rows(a,b,c,d,e,f,g,h,i,j,k,l) %>%
  drop_na()
```

```{r}
#binning data into different temperature zones
regions <- gbif_data %>% 
  mutate(thermal_zone = case_when(between(decimalLatitude, -20, 20.9) ~ "tropical", 
                                  between(decimalLatitude, 21, 40.9) ~ "subtropical",
                                  between(decimalLatitude, -21, -40.9) ~ "subtropical",
                                  between(decimalLatitude, -41, -60.9) ~ "temperate",
                                  between(decimalLatitude, 41, 60.9) ~ "temperate", 
                                  between(decimalLatitude, 61, 90) ~ "arctic", 
                                  between(decimalLatitude, -61, -90) ~ "arctic")) %>%
  mutate(genus = case_when(species == "Caiman latirostris" ~ "Caiman", 
         species %in% c("Drosophila birchii", "Drosophila melanogaster", 
                        "Drosophila simulans") ~ "Drosophila", 
         species == "Escherichia coli " ~ "Escherichia", 
         species == "Gryllus firmus " ~ "Gryllus",
         species %in% c("Limnodynastes peronii", 
                        "Limnodynastes tasmaniensis") ~ "Limnodynastes",
         species == "Mauremys reevesii" ~ "Mauremys",
         species == "Plestiodon chinensis " ~ "Plestiodon",
         species == "Rhodnius prolixus" ~ "Rhodnius",
         species == "Xylotrechus arvicola" ~ "Xylotrechus")) %>%
  drop_na()
```

@@

```{r}
#generating data for map overlay
world <- ne_countries(scale = "medium", returnclass = "sf")
#plotting up data 
ggplot(data = world)+
  geom_sf()+
  geom_point(data = regions, aes(x=decimalLongitude, y=decimalLatitude, color=genus), alpha = 0.5)+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -20, ymax = 20, alpha = 0.2, fill = "lightcoral")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 20, ymax = 40, alpha = 0.2, fill = "lightgoldenrod")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -40, ymax = -20, alpha = 0.2, fill = "lightgoldenrod")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 40, ymax = 60, alpha = 0.2, fill = "yellowgreen")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -60, ymax = -40, alpha = 0.2, fill = "yellowgreen")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 60, ymax = 90, alpha = 0.2, fill = "skyblue")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -90, ymax = -60, alpha = 0.2, fill = "skyblue")+
  geom_hline(yintercept = 0, alpha =0.5, linetype= "dashed")+
  theme_classic()+
  theme(legend.position="bottom")+
  ylab("Latitude")+
  xlab("Longitude")
```

```{r}
#writing to csv 
write_csv(regions, "gbif_regiondata.csv")
```

@@

>>>>

- S. enterica and E. coli had no data on GBIF and was therefore leftout.

@@

```{r}
#generating data for map overlay
regions <- read_csv("gbif_regiondata.csv")
world <- ne_countries(scale = "medium", returnclass = "sf")
#plotting up data 
ggplot(data = world)+
  geom_sf()+
  geom_point(data = regions, aes(x=decimalLongitude, y=decimalLatitude, color=genus), alpha = 0.5)+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -20, ymax = 20, alpha = 0.2, fill = "lightcoral")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 20, ymax = 40, alpha = 0.2, fill = "lightgoldenrod")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -40, ymax = -20, alpha = 0.2, fill = "lightgoldenrod")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 40, ymax = 60, alpha = 0.2, fill = "yellowgreen")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -60, ymax = -40, alpha = 0.2, fill = "yellowgreen")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 60, ymax = 90, alpha = 0.2, fill = "skyblue")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -90, ymax = -60, alpha = 0.2, fill = "skyblue")+
  geom_hline(yintercept = 0, alpha =0.5, linetype= "dashed")+
  theme_classic()+
  theme(legend.position="bottom")+
  ylab("Latitude")+
  xlab("Longitude")
```

```{r}
#attempting to pull thermal performance data from devRate

myDf <- exTropicalMoth$raw$egg
devRateModelAll(dfData = myDf)

devRateFind(species = "Drosophila melanogaster")
devRateFind(species = "Drosophila simulans")
```

@@

________

Additional notes/questions from figures last week

- Semenov 2007 and Kern et al 2015 were the two studies with the strangely smooth curves

  - Kern et al 2015--I changed the sample sizes but not sure they are right?
    - Looks less curved after sample size change
  -Not sure about the sample size for Semenov: sample size of 3 for 3 replicates, should I be multiplying that by the original CFU count?

- Removed Piccau et al (the big negative yi)--only study without a comparable constant and fluctuating mean

- How to best determine whether or not the temperatures each species is experiencing are heat stress temperatures?
  -It looks like in the final processes meta-analysis file there are quite a few species with range of temperatures

-Would it be useful to pull in and extract the thermal performance acclimation studies?
